
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=1in]{geometry}\usepackage{color,charter} \pagestyle{empty}
\usepackage[dvipsnames]{xcolor}

\begin{document}

\begin{itemize}
\item[\bf To:]  Professor Thomas Lee,  Editor\newline
{\em Journal of Computational and Graphical Statistics}
\item[\bf From:] Matthew Simpson, Jarad Niemi \& Vivekananda Roy, Iowa State University
\item[\bf Date:] \today
\item[\bf Re:] Submission of our paper to JCGS:
{\em ``Interweaving Markov Chain Monte Carlo Strategies for Efficient
    Estimation of Dynamic Linear Models''}
\end{itemize}

\subsection*{Comments for the Editor}

We hope that you find the enclosed article suitable for publication in The Journal of Computational and Graphical Statistics. This article applies the ideas of a previous article in JCGS by Yaming Yu and Xiao-Li Meng from 2011 which introduced interweaving as a strategy for constructing MCMC algorithms. We apply the ideas to dynamic linear models (DLMs), introduce a result that shows how difficult it is to implement their ASIS algorithm in DLMs, and construct several novel data augmentations for use in a general class of DLMs. We also assess the various available MCMC algorithms, interweaving and otherwise, in a simulation study.

There are several great candidates to review this paper. Either of Yaming Yu (yamingy@uci.edu, (949) 824-7361) and Xiao-Li Meng (meng@stat.harvard.edu, 617) 495-1603) would be a great reviewer since they came up with interweaving algorithms. Gregor Kastner (gregor.kastner@wu.ac.at, +43-1-313 36-5593) and Sylvia Fr{\"u}whirth-Schnatter (sylvia.fruehwirth-schnatter@wu.ac.at, +43-1-313 36-5581) are authors of the only paper in time series literature that we know of which implements an interweaving algorithm. Siddhartha Chib (chib@wustl.edu, (314) 935-4657) is another potential reviewer who is active in the MCMC for time series models literature.

Each author made substantial contributions to the article. Matthew Simpson is the corresponding author and can be reached at themattsimpson@gmail.com.

\end{document} 

